---
title: "multisemimarkov package"
author: "Yongho Lim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pgf,tikz,pgfplots}
  - \pgfplotsset{compat=1.15}
  - \usepackage{mathrsfs}
  - \usetikzlibrary{arrows}
  - \usetikzlibrary{patterns}
  - \usetikzlibrary{positioning}
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# multisemimarkov

<!-- badges: start -->

<!-- badges: end -->

Provides functionality with Multistate semi-Markov model with masked causes

## Installation

You can install the development version of `multisemimarkov` like so:

``` r
install.packages("multisemimarkov")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(multisemimarkov)
## basic example code
```

Multi-state model with partially latent cured and not cured states

```{=tex}
\centering
\begin{tikzpicture}
\node[draw, rectangle, rounded corners, text centered] (treatment) {Treatment (1)};
\node[draw, rectangle, above right= 1.5 of treatment, rounded corners, text centered] (notcured) {Not Cured};
\node[draw, rectangle, right=1.5 of notcured, rounded corners, text centered] (recurrence) {Recurrence (2)};
\node[draw, rectangle, below=4 of recurrence, rounded corners, text centered] (cancerdeath) {Cancer Death (3)};
\node[draw, rectangle,  below=4.1 of notcured, rounded corners, text centered] (cured) {Cured};
\draw[->, line width= 1, dashed] (treatment) --  node[above,font=\scriptsize]{}(notcured);
\draw[->, line width= 1, dashed] (treatment) --  node[above,font=\scriptsize]{}(cured);
\draw [->, line width= 1] (notcured) -- node[above,font=\scriptsize]{}(recurrence);
\draw [->, line width= 1] (recurrence) -- node[above,font=\scriptsize]{}(cancerdeath);
\draw [->, line width= 1] (notcured) -- node[above,font=\scriptsize]{}(cancerdeath);
\end{tikzpicture}
```
Multi-State Model Structure with Death Due to Other Causes State

```{=tex}
\centering
\begin{tikzpicture}
% nodes %
\node[draw, rectangle, rounded corners, text centered] (treatment) {Treatment (1)};
\node[draw, rectangle, above right= 1.5 of treatment, rounded corners, text centered] (notcured) {Not Cured};
\node[draw, rectangle, right=1.5 of notcured, rounded corners, text centered] (recurrence) {Recurrence (2)};
\node[draw, rectangle, below=2 of recurrence, rounded corners, text centered] (cancerdeath) {Cancer Death (3)};
\node[draw, rectangle,  below=5 of notcured, rounded corners, text centered] (cured) {Cured};
\node[draw, rectangle, right=0.6 of cured, rounded corners, text centered] (otherdeath) {Death due to Other Causes (4)};
% edges %
\draw[->, line width= 1, dashed] (treatment) --  node[above,font=\scriptsize]{}(notcured);
\draw[->, line width= 1, dashed] (treatment) --  node[above,font=\scriptsize]{}(cured);
\draw [->, line width= 1] (notcured) -- node[above,font=\scriptsize]{}(recurrence);
\draw [->, line width= 1] (recurrence) -- node[above,font=\scriptsize]{}(cancerdeath);
\draw [->, line width= 1] (notcured) -- node[above,font=\scriptsize]{}(cancerdeath);
\draw [->, line width= 1, dashed] (notcured) -- node[above,font=\scriptsize]{}(otherdeath);
\draw [->, line width= 1, dashed] (cured) -- node[above,font=\scriptsize]{}(otherdeath);
\end{tikzpicture}
```
